Standardization of ceramic shape: A case studay from the Iron Age pottery from northeastern Taiwan

### 2020-07-28 analysis/paper/paper.Rmd
change normalization scale

As a standardized measure of the spread of data, coefficients of variation (CV) allows us to directly compare variation in samples measured with different units or means. This is useful to compare the degree of standardization for archaeological assembles and enables comparison of variation across different sample sizes [@Eerkens2001, pp.498]. Following @Eerkens2001 and @Roux2003, we take this as our measurement of standardization in vessel shape variables: lower CV values reflect higher standardization, and thus increased craft specialization in the community. Given that CV is robust for positive values due to the representation by ratio, we normalized PC scores to a range between 0 and 1 before computation of CV.

@@

To compare pottery shape standardization across the three phases we investigate the distributions of the first three PC scores, taking the PC scores as proxy variables for vessel shape (Figure \@ref(fig:PCs-distribution)). The first PC, capturing vessel height and roundness, shows higher variation in the pre-European period compared to the Chinese period. That indicates that shape standardization was higher in the Chinese period compared to the pre-European period. The second PC also presents a similar pattern of higher shape standardization in the Chinese period compared to the other two phases.

@@

# function for normalized data
normalize <- function(x) (x-min(x))/(max(x)-min(x))

@@

ggplot(pc_plot,
       aes(phase,
           pc)) +
  geom_violin() + 
  ggforce::geom_sina() +
  xlab("") +
  facet_wrap(~ pcn, ncol = 1) +
  geom_text(data = cv_pcs_all,
            mapping = aes(x = c(2.1, 1.1, 3.1, 2.1, 1.1, 3.1, 2.1, 1.1, 3.1), 
                          y = 0.8, 
                          label = cv_label),
            size = 3,
            color= "red",
            hjust = -0.1,
            vjust= -1)+
  theme_minimal()

@@

@@

```{r size-analysis}

library(ggpubr)
library(geomorph)

>>>>

As a standardized measure of the spread of data, coefficients of variation (CV) allows us to directly compare variation in samples measured with different units or means. This is useful to compare the degree of standardization for archaeological assembles and enables comparison of variation across different sample sizes [@Eerkens2001, pp.498]. Following @Eerkens2001 and @Roux2003, we take this as our measurement of standardization in vessel shape variables: lower CV values reflect higher standardization, and thus increased craft specialization in the community. Given that CV is robust for positive values due to the representation by ratio, we normalized PC scores to a range between 1 and 2 before computation of CV.

@@

To compare pottery shape standardization across the three phases we investigate the distributions of the first three PC scores, taking the PC scores as proxy variables for vessel shape (Figure \@ref(fig:PCs-distribution)). The first PC, capturing vessel height and roundness, shows higher variation in the pre-European period compared to the Chinese period. That indicates that shape standardization was higher in the Chinese period compared to the pre-European period. The second PC also presents a similar pattern of higher shape standardization in the Chinese period compared to the other two phases. The CVs calculated of the three PC also support a general trend toward a more standardized shape over time, especially for first PC.

@@

# function for normalized data 
# normalize <- function(x) (x-min(x))/(max(x)-min(x)) # between 0-1
normalize <- function(x) (1 + ((x-min(x))* 2))/(max(x)-min(x))

@@

ggplot(pc_plot,
       aes(phase,
           pc)) +
  geom_violin() + 
  ggforce::geom_sina() +
  xlab("") +
  facet_wrap(~ pcn, ncol = 1) +
  geom_text(data = cv_pcs_all,
            mapping = aes(x = c(2.1, 1.1, 3.1, 2.1, 1.1, 3.1, 2.1, 1.1, 3.1), 
                          y = 2.1, 
                          label = cv_label),
            size = 3,
            color= "red",
            hjust = -0.1,
            vjust= -1)+
  theme_minimal()

@@

pc2_chinese <-
  pc2 %>% 
  filter(y == "Chinese") %>% 
  mutate(cv = round(sd(x)/ mean(x), digits = 2)*100)

@@

```{r size-analysis}
library(ggpubr)
library(geomorph)
